---
title: "MAIN analysis"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(brms)
library(ggthemes)
```

# Read data

```{r}
data <- read_csv("../data/data.csv")
```
## Descriptives
```{r}
n_distinct(data$child)
# [1] 1405

n_distinct(data$testlanguage)
#[1] 33

range(data$age)/12
#3.416667 13.833333

data%>%group_by(testlanguage)%>%summarise(n = n_distinct(child), n = 33)
#   testlanguage     n
#    <chr>        <int>
#  1 aka             39
#  2 eng            296
#  3 est             27
#  4 fin             55
#  5 geo             20
#  6 ger             90
#  7 ghc             15
#  8 gle              6
#  9 gon             75
# 10 heb             44
# 11 hin            121
# 12 hlb             32
# 13 hrv             91
# 14 ita             87
# 15 kan             18
# 16 kaz             14
# 17 lit             28
# 18 lug             24
# 19 nld            127
# 20 per             81
# 21 pol             83
# 22 rus            237
# 23 sah             30
# 24 sgh             20
# 25 slk             62
# 26 swe            150
# 27 tgk             16
# 28 tgl             18
# 29 tur             45
# 30 uig              5
# 31 urd             24
# 32 uzb             19
# 33 yue             24
```


```{r}
# scale age for modelling
mdata <- data%>%
  mutate(age = scale(age))
```

# 1. What are the differences in children’s overall narrative skills across languages?

```{r}
# aggregate data per individual for binomial model
mdata1 <- mdata%>%
  group_by(child, age, mode, testlanguage, ling)%>%
  summarise(sum = sum(response),
            n = length(response))
```


```{r}
bm1 <- brm(sum|trials(n) ~ age * ling + mode + (1|child) + (age | testlanguage), 
    data = mdata1,
    chains = 4, 
    cores = 4,
    iter = 4000,
    family = binomial(),
    threads = threading(8), #to speed things up, comment out if not on a cluster
    backend = "cmdstanr",
    control = list(adapt_delta = 0.95, max_treedepth = 20)
    )

saveRDS(bm1, "../saves/bm1.rds")
```


```{r}
bm1_1 <- brm(sum|trials(n) ~ age + ling + mode + (1|child) + (age | testlanguage), 
    data = mdata1,
    chains = 4, 
    cores = 4,
    iter = 4000,
    family = binomial(),
    threads = threading(8), #to speed things up, comment out if not on a cluster
    backend = "cmdstanr",
    control = list(adapt_delta = 0.95, max_treedepth = 20)
    )

saveRDS(bm1, "../saves/bm1_1.rds")
```

## Fixed effects

```{r}
fixef(bm1)
```
                      Estimate  Est.Error       Q2.5       Q97.5
Intercept            0.3275707 0.07727257  0.1742991  0.48063550
age                  0.6678299 0.10240489  0.4666000  0.87303440
lingMonolingual     -0.1313758 0.09087711 -0.3085195  0.04677174
modetelling         -0.3652205 0.02431607 -0.4122890 -0.31800940
age:lingMonolingual -0.4463749 0.12358384 -0.6868958 -0.20327683


```{r}
fixef(bm1_1)
```

## Visualization

```{r}
ndata1 <- mdata1%>%
  select(testlanguage, age, ling)%>%
  group_by(testlanguage, ling)%>%
  distinct(age)%>%
  expand_grid(mode = unique(mdata$mode))%>%
  mutate(n = 1)

post1 <- fitted(bm1_1, newdata = ndata1, re_formula =~(age | testlanguage))%>%
  as_tibble() %>%
  bind_cols(ndata1)%>%
  mutate(age = age * sd(data$age) + mean(data$age))%>%
  mutate(mode = str_to_title(mode))%>%
  mutate(mode = factor(mode, levels = c("Telling","Retelling")))

ndata1_2 <- mdata%>%
  select(age, ling)%>%
  group_by(ling)%>%
  distinct(age)%>%
  expand_grid(mode = unique(mdata$mode))%>%
  mutate(n = 1)

post1_2 <- fitted(bm1_1, newdata = ndata1_2, re_formula = NA)%>%
  as_tibble() %>%
  bind_cols(ndata1_2)%>%
  mutate(age = age * sd(data$age) + mean(data$age))%>%
  mutate(mode = str_to_title(mode))%>%
  mutate(mode = factor(mode, levels = c("Telling","Retelling")))


saveRDS(post1, "../saves/post1.rds")
saveRDS(post1_2, "../saves/post1_2.rds")


post1 <- readRDS("../saves/post1.rds")
post1_2 <- readRDS("../saves/post1_2.rds")


ggplot()+
  geom_line(data = post1, aes(x = age/12, y = Estimate, ymin = Q2.5, ymax = Q97.5, col = ling, group = interaction(ling,testlanguage)), stat = "identity", alpha = .25)+
  geom_smooth(data = post1_2, aes(x = age/12, y = Estimate, ymin = Q2.5, ymax = Q97.5, fill =ling, col = ling), stat = "identity", alpha = .25)+
  theme_bw()+
  scale_color_ptol(name = "Language status")+
  scale_fill_ptol(name = "Language status")+
  ylim(0,1)+
  facet_wrap(~mode)+
  labs(x = "Age in years", y="Probability of producing a narrative component")+
  theme(legend.position = c(0.8, 0.15), legend.background = element_rect(colour = "black", size = 0.5, linetype = "solid"))
```

```{r}
ggsave("../visuals/fig2.png", width = 10, height = 5, scale = 1)
```

# 2. How does the development of bilinguals’ narrative skills in L1 compare to that in L2? 

```{r}
mdata2 <- mdata%>%
  filter(ling == "Bilingual")%>%
  group_by(child, age, mode, testlanguage, test_lang_status)%>%
  summarise(sum = sum(response),
            n = length(response))


bm2 <- brm(sum|trials(n) ~ age + test_lang_status + mode + (1|child) + (age | testlanguage), 
    data = mdata2,
    chains = 4, 
    cores = 4,
    iter = 4000,
    family = binomial(),
    threads = threading(8), #to speed things up, comment out if not on a cluster
    backend = "cmdstanr",
    control = list(adapt_delta = 0.9, max_treedepth = 20)
    )%>%add_criterion("loo") #to speed things up, comment out if not on a cluster)


saveRDS(bm2, "../saves/bm2.rds")

bm2_1 <- brm(sum|trials(n) ~ age * test_lang_status + mode + (1|child) + (age | testlanguage), 
    data = mdata2,
    chains = 4, 
    cores = 4,
    iter = 4000,
    family = binomial(),
    threads = threading(8), #to speed things up, comment out if not on a cluster
    backend = "cmdstanr",
    control = list(adapt_delta = 0.9, max_treedepth = 20)
    )%>%add_criterion("loo") #to speed things up, comment out if not on a cluster)


saveRDS(bm2_1, "../saves/bm2_1.rds")
```

## Model comparison

```{r}
print(loo_compare(bm2, bm2_1)%>%as_tibble(rownames = "model"))
```
  model elpd_diff  se_diff    elpd_loo   se_elpd_loo p_loo   se_p_loo looic se_looic
  <chr> <compar.l> <compar.l> <compar.l> <compar.l>  <compa> <compar> <com> <compar>
1 bm2    0.0000000 0.000000   -5169.621  38.54474    513.91… 13.95926 1033… 77.08947
2 bm2_1 -0.5349237 0.983548   -5170.155  38.51985    515.14… 13.93043 1034… 77.03971


## Fixed effects

```{r}
fixef(bm2)
```

                     Estimate  Est.Error        Q2.5      Q97.5
Intercept           0.1886201 0.08289399  0.02434448  0.3499550
age                 0.5779004 0.10682822  0.36792365  0.7907588
test_lang_statusL2  0.2192306 0.05867445  0.10728780  0.3349802
modetelling        -0.3584498 0.02497057 -0.40760430 -0.3101704

## Visualization

```{r}
ndata2 <- mdata2%>%
  select(testlanguage,test_lang_status, age, mode)%>%
  group_by(testlanguage,test_lang_status, mode)%>%
  distinct(age)%>%
  mutate(n = 1)


post2 <- fitted(bm2, newdata = ndata2, re_formula =~(age | testlanguage))%>%
  as_tibble() %>%
  bind_cols(ndata2)%>%
  mutate(age = age * sd(data$age) + mean(data$age))%>%
  mutate(mode = str_to_title(mode))%>%
  mutate(mode = factor(mode, levels = c("Telling","Retelling")))


ndata2_2 <- mdata2%>%
  ungroup()%>%
  distinct(mode, test_lang_status)%>%
  expand_grid(age = unique(mdata2$age))%>%
  mutate(n = 1)

post2_2 <- fitted(bm2, newdata = ndata2_2, re_formula = NA)%>%
  as_tibble() %>%
  bind_cols(ndata2_2)%>%
  mutate(age = age * sd(data$age) + mean(data$age))%>%
  mutate(mode = str_to_title(mode))%>%
  mutate(mode = factor(mode, levels = c("Telling","Retelling")))


saveRDS(post2, "../saves/post2.rds")
saveRDS(post2_2, "../saves/post2_2.rds")


post2 <- readRDS("../saves/post2.rds")
post2_2 <- readRDS("../saves/post2_2.rds")


ggplot()+
  geom_line(data = post2, aes(x = age/12, y = Estimate, ymin = Q2.5, ymax = Q97.5, col = test_lang_status, group = interaction(test_lang_status,testlanguage)), stat = "identity", alpha = .25)+
  geom_smooth(data = post2_2, aes(x = age/12, y = Estimate, ymin = Q2.5, ymax = Q97.5, fill =test_lang_status, col = test_lang_status), stat = "identity", alpha = .25)+
  theme_bw()+
  scale_color_manual(name = "Test language", values = c("#345c83", "#40531B"))+
  scale_fill_manual(name = "Test language", values = c("#345c83", "#40531B"))+
  ylim(0,1)+
  facet_grid(~mode)+
  labs(x = "Age", y="Probability of producing a narrative component")
```

```{r}
ggsave("../visuals/fig3.png", width = 10, height = 5, scale = 1)
```

# 3. What is the general pattern in children’s use of factual and inferred components across the investigated languages?

run script "run_bm3.R"

```{r}
bm3 <- readRDS("../saves/bm3.rds")
```

## Fixed effects

```{r}
fixef(bm3)
```

                      Estimate  Est.Error       Q2.5       Q97.5
Intercept            0.5736588 0.66094813 -0.8164857  1.87133975
age                  0.7904910 0.28360935  0.2838487  1.33117925
lingMonolingual     -0.1295025 0.10735855 -0.3360821  0.08074064
modetelling         -0.4118338 0.02692149 -0.4640448 -0.35936258
age:lingMonolingual -0.5162402 0.14297884 -0.7930640 -0.23489268


## Item effects

```{r}
ranef(bm3)$item
```


          Estimate Est.Error       Q2.5     Q97.5
attempt  0.8244836 0.6541805 -0.4765082 2.2126055
goal    -0.8948328 0.6547706 -2.1964730 0.4923131
ist     -0.5591364 0.6529128 -1.8568003 0.8227929
outcome  0.8138825 0.6530311 -0.4749704 2.2059632

## Visualization

### Probability of component by language for avergae age
```{r}
ndata3 <- mdata%>%
  distinct(item)%>%
  expand_grid(testlanguage = unique(mdata$testlanguage))%>%
  mutate(mode = "telling", 
         age = 0, 
         ling = "Bilingual")


post3 <- fitted(bm3, newdata = ndata3, re_formula =~(testlanguage | item))%>%
  as_tibble() %>%
  bind_cols(ndata3)%>%
  mutate(testlanguage = recode(testlanguage,
                           "eng" = "English", "per" = "Persian", "hrv" = "Croatian", 
                           "rus" = "Russian", "est" = "Estonian", "fin" = "Finnish", 
                           "swe" = "Swedish", "geo" = "Georgian", "ger" = "German", 
                           "tur" = "Turkish", "aka" = "Akan", "urd" = "Urdu", 
                           "yue" = "Cantonese", "hin" = "Hindi", "kan" = "Kannada", 
                           "gon" = "Gondi", "hlb" = "Halbi", "gle" = "Irish", 
                           "heb" = "Hebrew", "ita" = "Italian", "kaz" = "Kazakh", 
                           "uig" = "Uyghur", "uzb" = "Uzbek", "lit" = "Lithuanian", 
                           "nld" = "Dutch", "tgl" = "Tagalog", "pol" = "Polish", 
                           "ghc" = "Gaelic", "slk" = "Slovak", "sgh" = "Shughni", 
                           "tgk" = "Tajik", "lug" = "Luganda", "sah" = "Yakut"))%>%
  filter(item != "setting")%>%
  mutate(item = recode(item,
                        "ist" = "Internal State", 
                       "goal" = "Goal",
                       "attempt" = "Attempt",
                       "outcome" = "Outcome"))%>%
  mutate(item = factor(item, levels = c("Internal State", "Goal", "Attempt", "Outcome")))


saveRDS(post3, "../saves/post3.rds")

post3 <- readRDS("../saves/post3.rds")

ggplot()+
  geom_pointrange(data = post3, aes(y = reorder(testlanguage, Estimate), x = Estimate, xmin = Q2.5, xmax = Q97.5, col = item), alpha = .75, position = position_dodge(width = 0.5))+
  theme_bw()+
  scale_color_viridis_d(name = "Component")+
  labs(y = "Language", x="Probability to produce the component")

```

```{r}
ggsave("../visuals/fig4.png", height = 8, width = 12, scale = 1)
```

### Probability of component by age across languages

```{r}
ndata3_1 <- mdata%>%
  distinct(item, age, ling)%>%
  mutate(mode = "telling")

post3_1 <- fitted(bm3, newdata = ndata3_1, re_formula =~(age | item), )%>%
  as_tibble() %>%
  bind_cols(ndata3_1)%>%
  mutate()%>%
  filter(item != "setting")%>%
  mutate(item = recode(item,
                        "ist" = "Internal State", 
                       "goal" = "Goal",
                       "attempt" = "Attempt",
                       "outcome" = "Outcome"))%>%
  mutate(item = factor(item, levels = c("Internal State", "Goal", "Attempt", "Outcome")))%>%
  mutate(age = age * sd(data$age) + mean(data$age))%>%
  mutate(mode = str_to_title(mode))

saveRDS(post3_1, "../saves/post3_1.rds")

post3_1 <- readRDS("../saves/post3_1.rds")

ggplot()+
  geom_smooth(data = post3_1, aes(x = age/12, y = Estimate, ymin = Q2.5, ymax = Q97.5, fill =item, col = item), stat = "identity", alpha = .25)+
  theme_bw()+
  scale_color_viridis_d(name = "Component")+
  scale_fill_viridis_d(name = "Component")+
  facet_grid(~ling)+
  labs(y = "Probability to produce the component", x="Age")+
  theme(legend.position = c(0.9, 0.2))
```

```{r}
ggsave("../visuals/fig5.png", width = 10, height = 5, scale = 1)
```

# 4. How do complete episodes evolve across languages and in monolingual vs bilingual children?

```{r}
# read in data coded for complete narratives (attempt, goal or internal state, and outcome)
nar_data1 <- read_csv("../data/nar_data1.csv")
```

```{r}
mnar_data1 <- nar_data1%>%
  mutate(age = scale(age))%>%
  group_by(child, age, mode, testlanguage, ling)%>%
  summarise(sum = sum(complete),
            n = length(complete))

bm4 <- brm(sum|trials(n) ~ age + ling + mode + (1|child) + (age | testlanguage), 
    data = mnar_data1,
    chains = 4, 
    cores = 4,
    iter = 4000,
    family = binomial(),
    threads = threading(8), #to speed things up, comment out if not on a cluster
    backend = "cmdstanr",
    control = list(adapt_delta = 0.95, max_treedepth = 20)
    )%>%add_criterion("loo")  #to speed things up, comment out if not on a cluster)


saveRDS(bm4, "../saves/bm4.rds")

bm4_1 <- brm(sum|trials(n) ~ age * ling + mode + (1|child) + (age | testlanguage), 
    data = mnar_data1,
    chains = 4, 
    cores = 4,
    iter = 4000,
    family = binomial(),
    threads = threading(8), #to speed things up, comment out if not on a cluster
    backend = "cmdstanr",
    control = list(adapt_delta = 0.95, max_treedepth = 20)
    )%>%add_criterion("loo") 

saveRDS(bm4_1, "../saves/bm4_1.rds")

```

## Model comparison

```{r}
print(loo_compare(bm4, bm4_1))
```

    elpd_diff se_diff
bm4_1  0.0       0.0   
bm4   -1.5       1.7 

## Fixed effects

### Model wo interaction

```{r}
fixef(bm4)
```

                Estimate  Est.Error       Q2.5      Q97.5
Intercept       -0.4060002 0.15710090 -0.7241534 -0.1043293
age              0.9367130 0.16534866  0.6220069  1.2682790
lingMonolingual  0.3113266 0.10762288  0.1011727  0.5218090
modetelling     -0.4780694 0.06066418 -0.5984158 -0.3549043

### Interaction model 

```{r}
fixef(bm4_1)
```

                       Estimate  Est.Error       Q2.5       Q97.5
Intercept           -0.36217769 0.16095835 -0.6776961 -0.04212305
age                  1.02556981 0.18166173  0.6822135  1.39046625
lingMonolingual      0.02123661 0.20326860 -0.3751446  0.41633225
modetelling         -0.47433415 0.06030159 -0.5920365 -0.35511398
age:lingMonolingual -0.46447991 0.27314590 -1.0145452  0.06399581

### Visualization

```{r}
ndata4 <- mnar_data1%>%
  ungroup()%>%
  distinct(testlanguage)%>%
  expand_grid(ling = unique(nar_mdata$ling))%>%
  mutate(mode = "telling", 
         age = 0, 
         n = 1)


post4 <- fitted(bm4, newdata = ndata4, re_formula =~(1 | testlanguage))%>%
  as_tibble() %>%
  bind_cols(ndata4)%>%
  mutate(testlanguage = recode(testlanguage,
                           "eng" = "English", "per" = "Persian", "hrv" = "Croatian", 
                           "rus" = "Russian", "est" = "Estonian", "fin" = "Finnish", 
                           "swe" = "Swedish", "geo" = "Georgian", "ger" = "German", 
                           "tur" = "Turkish", "aka" = "Akan", "urd" = "Urdu", 
                           "yue" = "Cantonese", "hin" = "Hindi", "kan" = "Kannada", 
                           "gon" = "Gondi", "hlb" = "Halbi", "gle" = "Irish", 
                           "heb" = "Hebrew", "ita" = "Italian", "kaz" = "Kazakh", 
                           "uig" = "Uyghur", "uzb" = "Uzbek", "lit" = "Lithuanian", 
                           "nld" = "Dutch", "tgl" = "Tagalog", "pol" = "Polish", 
                           "ghc" = "Gaelic", "slk" = "Slovak", "sgh" = "Shughni", 
                           "tgk" = "Tajik", "lug" = "Luganda", "sah" = "Yakut"))


saveRDS(post4, "../saves/post4.rds")

post4 <- readRDS("../saves/post4.rds")


ggplot()+
  geom_pointrange(data = post4, aes(y = testlanguage, x = Estimate, xmin = Q2.5, xmax = Q97.5, col = ling), alpha = .75, position = position_dodge(width = 0.3))+
  theme_bw()+
  scale_color_ptol(name = "Language status")+
  labs(y = "Language", x="Probability of producing complete narrative")
```

```{r}
ggsave("../visuals/fig6.png", height = 8, width = 12, scale = 1)
```

# 4.1. Complete narratives in bilinguals

```{r}
nar_data1_bi <- nar_data1%>%
  filter(ling == "Bilingual")%>%
  mutate(age = scale(age))%>%
  group_by(child, age, mode, testlanguage, test_lang_status)%>%
  summarise(sum = sum(complete),
            n = length(complete))

bm4_bi <- brm(sum|trials(n) ~ age + test_lang_status + mode + (1|child) + (age | testlanguage), 
    data = nar_data1_bi,
    chains = 4, 
    cores = 4,
    iter = 4000,
    family = binomial(),
    threads = threading(8), #to speed things up, comment out if not on a cluster
    backend = "cmdstanr",
    control = list(adapt_delta = 0.95, max_treedepth = 20)
    )%>%add_criterion("loo") 

saveRDS(bm4_bi, "../saves/bm4_bi.rds")

```

## Fixed effects

```{r}
fixef(bm4_bi)
```

                      Estimate  Est.Error       Q2.5       Q97.5
Intercept          -0.24477052 0.17858580 -0.6019724  0.09200422
age                 0.97142966 0.19570141  0.5981369  1.36177025
test_lang_statusL2  0.03601371 0.13283665 -0.2135332  0.30476942
modetelling        -0.46599688 0.06362462 -0.5926337 -0.34166000

# 4.2. Complete narratives ignoring internal states (narrative = goal + attempt + outcome)

```{r}
# read in data coded for complete narratives (attempt, goal, and outcome)
nar_data2 <- read_csv("../data/nar_data2.csv")
```

```{r}
mnar_data2 <- nar_data2%>%
  mutate(age = scale(age))%>%
  group_by(child, age, mode, testlanguage, ling)%>%
  summarise(sum = sum(complete),
            n = length(complete))

bm4_2 <- brm(sum|trials(n) ~ age + ling + mode + (1|child) + (age | testlanguage), 
    data = mnar_data2,
    chains = 4, 
    cores = 4,
    iter = 4000,
    family = binomial(),
    threads = threading(8), #to speed things up, comment out if not on a cluster
    backend = "cmdstanr",
    control = list(adapt_delta = 0.95, max_treedepth = 20)
    )%>%add_criterion("loo")  #to speed things up, comment out if not on a cluster)


saveRDS(bm4_2, "../saves/bm4_2.rds")
```

## Fixed effects

```{r}
fixef(bm4_2)
```

                  Estimate  Est.Error         Q2.5      Q97.5
Intercept       -1.4929088 0.18578886 -1.864353000 -1.1280023
age              0.8041994 0.19514296  0.428887475  1.2097660
lingMonolingual  0.2862603 0.14310121  0.002662163  0.5656250
modetelling     -0.6060893 0.07683911 -0.755128450 -0.4558087



