---
title: "MAIN analysis"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(brms)
library(ggthemes)
```

# Read data

```{r}
data <- read_csv("../data/data.csv")
```
## Descriptives
```{r}
n_distinct(data$child)
# [1] 2740

n_distinct(data$testlanguage)
#[1] 44

range(data$age)/12
#3 14

sd(data$age)/12
#2.335605

mean(data$age)/12
#7.342638

print(data%>%group_by(testlanguage)%>%summarise(n = n_distinct(child)),n=44)
#    testlanguage     n
#    <chr>        <int>
#  1 afr            110
#  2 aka             39
#  3 cat             70
#  4 cmn             19
#  5 cor             55
#  6 dan             72
#  7 ell             48
#  8 eng            391
#  9 esp             70
# 10 est             27
# 11 fin             55
# 12 geo             20
# 13 ger            205
# 14 ghc             15
# 15 gle             35
# 16 gon             75
# 17 heb             44
# 18 hin            121
# 19 hlb             32
# 20 hrv             91
# 21 ita            142
# 22 kan             18
# 23 kaz             14
# 24 kor            522
# 25 lit             28
# 26 lug             24
# 27 msa             61
# 28 nld            127
# 29 per             81
# 30 pol             83
# 31 rus            288
# 32 sah             30
# 33 sgh             20
# 34 slk            118
# 35 swe            150
# 36 tgk             16
# 37 tgl             18
# 38 tur             45
# 39 uig              5
# 40 ukr             19
# 41 urd             24
# 42 uzb             19
# 43 xho            111
# 44 yue             24
```

```{r}
# scale age for modelling
mdata <- data%>%
  mutate(age = scale(age))
```

# 0. Overall development of narrative skills

```{r}
# aggregate data per individual for binomial model
mdata0 <- mdata%>%
  group_by(child, age, testlanguage)%>%
  summarise(sum = sum(response),
            n = length(response))
```

```{r}
# bm0 <- brm(sum|trials(n) ~ age + (1|child) + (age | testlanguage), 
#     data = mdata0,
#     chains = 4, 
#     cores = 4,
#     iter = 4000,
#     family = binomial(),
#     threads = threading(8), #to speed things up, comment out if not on a cluster
#     backend = "cmdstanr",
#     control = list(adapt_delta = 0.95, max_treedepth = 20)
#     )%>%add_criterion("loo")
# 
# saveRDS(bm0, "../saves/bm0.rds")
```

## Fixed effects

```{r}
fixef(bm0)
```
            Estimate  Est.Error       Q2.5     Q97.5
Intercept 0.02570466 0.06510717 -0.1023787 0.1556279
age       0.59404431 0.08762003  0.4278555 0.7686384

## Visualization

```{r}
# ndata0 <- mdata0%>%
#   select(testlanguage, age)%>%
#   group_by(testlanguage)%>%
#   distinct(age)%>%
#   mutate(n = 1)
# 
# post0 <- fitted(bm0, newdata = ndata0, re_formula =~(age | testlanguage))%>%
#   as_tibble() %>%
#   bind_cols(ndata0)%>%
#   mutate(age = age * sd(data$age) + mean(data$age))
# 
# ndata0_2 <- mdata%>%
#   distinct(age)%>%
#   mutate(n = 1)
# 
# post0_2 <- fitted(bm0, newdata = ndata0_2, re_formula = NA)%>%
#   as_tibble() %>%
#   bind_cols(ndata0_2)%>%
#   mutate(age = age * sd(data$age) + mean(data$age))
# 
# 
# saveRDS(post0, "../saves/post0.rds")
# saveRDS(post0_2, "../saves/post0_2.rds")


post0 <- readRDS("../saves/post0.rds")
post0_2 <- readRDS("../saves/post0_2.rds")


ggplot()+
  geom_line(data = post0, aes(x = age/12, y = Estimate, ymin = Q2.5, ymax = Q97.5, group = testlanguage), stat = "identity", alpha = .25)+
  geom_smooth(data = post0_2, aes(x = age/12, y = Estimate, ymin = Q2.5, ymax = Q97.5), col = "black", stat = "identity", alpha = .25)+
  theme_bw()+
  ylim(0,1)+
  labs(x = "Age in years", y="Probability of producing a narrative component")+
  theme(legend.position = c(0.8, 0.15), legend.background = element_rect(colour = "black", size = 0.5, linetype = "solid"))
```
```{r}
ggsave("../visuals/fig2.png", width = 6, height = 4, scale = 1)
```

# 1. What are the differences in children’s overall narrative skills across languages?

```{r}
# aggregate data per individual for binomial model
mdata1 <- mdata%>%
  group_by(child, age, mode, testlanguage, ling)%>%
  summarise(sum = sum(response),
            n = length(response))
```


```{r}
# bm1 <- brm(sum|trials(n) ~ age * ling + mode + (1|child) + (age | testlanguage), 
#     data = mdata1,
#     chains = 4, 
#     cores = 4,
#     iter = 4000,
#     family = binomial(),
#     threads = threading(8), #to speed things up, comment out if not on a cluster
#     backend = "cmdstanr",
#     control = list(adapt_delta = 0.95, max_treedepth = 20)
#     )%>%add_criterion("loo")
# 
# saveRDS(bm1, "../saves/bm1.rds")
```


```{r}
# bm1_1 <- brm(sum|trials(n) ~ age + ling + mode + (1|child) + (age | testlanguage), 
#     data = mdata1,
#     chains = 4, 
#     cores = 4,
#     iter = 4000,
#     family = binomial(),
#     threads = threading(8), #to speed things up, comment out if not on a cluster
#     backend = "cmdstanr",
#     control = list(adapt_delta = 0.95, max_treedepth = 20)
#     )%>%add_criterion("loo")
# 
# saveRDS(bm1, "../saves/bm1_1.rds")
```


```{r}
# print(loo_compare(bm1, bm1_1)%>%as_tibble(rownames = "model"))
```
model elpd_diff  se_diff    elpd_loo   se_elpd_loo p_loo  se_p_loo looic
  <chr> <compar.l> <compar.l> <compar.l> <compar.l>  <comp> <compar> <com>
1 bm1    0.000000  0.00000    -11319.19  58.29841    1231.… 23.52328 2263…
2 bm1_1 -7.755914  3.92498    -11326.94  58.23827    1238.… 23.51072 2265…

## Fixed effects

```{r}
fixef(bm1)
```
                       Estimate  Est.Error       Q2.5       Q97.5
Intercept            0.37084631 0.07303083  0.2274967  0.51417952
age                  0.59967155 0.08867293  0.4244529  0.77787010
lingMonolingual     -0.06333196 0.04089941 -0.1426443  0.01715709
modetelling         -0.56449412 0.01777624 -0.5991887 -0.53009042
age:lingMonolingual -0.20925446 0.05522629 -0.3176186 -0.09968928


## Visualization

```{r}
# ndata1 <- mdata1%>%
#   select(testlanguage, age, ling)%>%
#   group_by(testlanguage, ling)%>%
#   distinct(age)%>%
#   expand_grid(mode = unique(mdata$mode))%>%
#   mutate(n = 1)
# 
# post1 <- fitted(bm1, newdata = ndata1, re_formula =~(age | testlanguage))%>%
#   as_tibble() %>%
#   bind_cols(ndata1)%>%
#   mutate(age = age * sd(data$age) + mean(data$age))%>%
#   mutate(mode = str_to_title(mode))%>%
#   mutate(mode = factor(mode, levels = c("Telling","Retelling")))
# 
# ndata1_2 <- mdata%>%
#   select(age, ling)%>%
#   group_by(ling)%>%
#   distinct(age)%>%
#   expand_grid(mode = unique(mdata$mode))%>%
#   mutate(n = 1)
# 
# post1_2 <- fitted(bm1, newdata = ndata1_2, re_formula = NA)%>%
#   as_tibble() %>%
#   bind_cols(ndata1_2)%>%
#   mutate(age = age * sd(data$age) + mean(data$age))%>%
#   mutate(mode = str_to_title(mode))%>%
#   mutate(mode = factor(mode, levels = c("Telling","Retelling")))
# 
# 
# saveRDS(post1, "../saves/post1.rds")
# saveRDS(post1_2, "../saves/post1_2.rds")


post1 <- readRDS("../saves/post1.rds")
post1_2 <- readRDS("../saves/post1_2.rds")


ggplot()+
  geom_line(data = post1, aes(x = age/12, y = Estimate, ymin = Q2.5, ymax = Q97.5, col = ling, group = interaction(ling,testlanguage)), stat = "identity", alpha = .25)+
  geom_smooth(data = post1_2, aes(x = age/12, y = Estimate, ymin = Q2.5, ymax = Q97.5, fill =ling, col = ling), stat = "identity", alpha = .25)+
  theme_bw()+
  scale_color_ptol(name = "Language status")+
  scale_fill_ptol(name = "Language status")+
  ylim(0,1)+
  facet_wrap(~mode)+
  labs(x = "Age in years", y="Probability of producing a narrative component")+
  theme(legend.position = c(0.8, 0.15), legend.background = element_rect(colour = "black", size = 0.5, linetype = "solid"))
```

```{r}
ggsave("../visuals/fig2.png", width = 10, height = 5, scale = 1)
```

# 1.1. Differences for languages with data from mono and bilinguals

```{r}
data_ling <- read_csv("../data/data_multi_ling.csv")
```

```{r}
mdata1_ling <- data_ling%>%
  mutate(age = scale(age))%>%
  group_by(child, age, mode, testlanguage, ling)%>%
  summarise(sum = sum(response),
            n = length(response))
```


```{r}
# bm1_2 <- brm(sum|trials(n) ~ age * ling + mode + (1|child) + (age | testlanguage), 
#     data = mdata1_ling,
#     chains = 4, 
#     cores = 4,
#     iter = 4000,
#     family = binomial(),
#     threads = threading(8), #to speed things up, comment out if not on a cluster
#     backend = "cmdstanr",
#     control = list(adapt_delta = 0.95, max_treedepth = 20)
#     )%>%add_criterion("loo")
# 
# saveRDS(bm1_2, "../saves/bm1_2.rds")
```

## Fixed effects

```{r}
# fixef(bm1_2)
```

                      Estimate  Est.Error        Q2.5       Q97.5
Intercept           -0.2594510 0.07600014 -0.41577172 -0.11193113
age                  0.4427069 0.07606136  0.29453743  0.59732642
lingMonolingual      0.1847095 0.04559862  0.09574817  0.27628672
modetelling         -0.1622292 0.05036460 -0.26252720 -0.06285328
age:lingMonolingual -0.2596995 0.04475827 -0.34868975 -0.17057035

## Visualization

```{r}
# ndata12 <- mdata1_ling%>%
#   select(testlanguage, age, ling)%>%
#   group_by(testlanguage, ling)%>%
#   distinct(age)%>%
#   expand_grid(mode = unique(data_ling$mode))%>%
#   mutate(n = 1)
# 
# post12 <- fitted(bm1_2, newdata = ndata12, re_formula =~(age | testlanguage))%>%
#   as_tibble() %>%
#   bind_cols(ndata12)%>%
#   mutate(age = age * sd(data_ling$age) + mean(data_ling$age))%>%
#   mutate(mode = str_to_title(mode))%>%
#   mutate(mode = factor(mode, levels = c("Telling","Retelling")))
# 
# ndata12_2 <- mdata1_ling%>%
#   select(age, ling)%>%
#   group_by(ling)%>%
#   distinct(age)%>%
#   expand_grid(mode = unique(data_ling$mode))%>%
#   mutate(n = 1)
# 
# post12_2 <- fitted(bm1_2, newdata = ndata12_2, re_formula = NA)%>%
#   as_tibble() %>%
#   bind_cols(ndata12_2)%>%
#   mutate(age = age * sd(data_ling$age) + mean(data_ling$age))%>%
#   mutate(mode = str_to_title(mode))%>%
#   mutate(mode = factor(mode, levels = c("Telling","Retelling")))
# 
# 
# saveRDS(post12, "../saves/post12.rds")
# saveRDS(post12_2, "../saves/post12_2.rds")


post12 <- readRDS("../saves/post12.rds")
post12_2 <- readRDS("../saves/post12_2.rds")


ggplot()+
  geom_line(data = post12, aes(x = age/12, y = Estimate, ymin = Q2.5, ymax = Q97.5, col = ling, group = interaction(ling,testlanguage)), stat = "identity", alpha = .25)+
  geom_smooth(data = post12_2, aes(x = age/12, y = Estimate, ymin = Q2.5, ymax = Q97.5, fill =ling, col = ling), stat = "identity", alpha = .25)+
  theme_bw()+
  scale_color_ptol(name = "Language status")+
  scale_fill_ptol(name = "Language status")+
  ylim(0,1)+
  facet_wrap(~mode)+
  labs(x = "Age in years", y="Probability of producing a narrative component")+
  theme(legend.position = c(0.8, 0.15), legend.background = element_rect(colour = "black", size = 0.5, linetype = "solid"))
```

```{r}
ggsave("../visuals/fig2.png", width = 10, height = 5, scale = 1)
```

## Joint Figure 1 and 1.1

```{r}

postl <- bind_rows(
  post1%>%mutate(type = "All data"),
  post12%>%mutate(type = "Languages with data\n for mono- and bilinguals")
) 


posta <- bind_rows(
  post1_2%>%mutate(type = "All data"),
  post12_2%>%mutate(type = "Languages with data\n for mono- and bilinguals")
) 

ggplot()+
  geom_line(data = postl, aes(x = age/12, y = Estimate, ymin = Q2.5, ymax = Q97.5, col = ling, group = interaction(ling,testlanguage)), stat = "identity", alpha = .25)+
  geom_smooth(data = posta, aes(x = age/12, y = Estimate, ymin = Q2.5, ymax = Q97.5, fill =ling, col = ling), stat = "identity", alpha = .25)+
  theme_bw()+
  scale_color_ptol(name = "Language status")+
  scale_fill_ptol(name = "Language status")+
  ylim(0,1)+
  facet_grid(type~mode)+
  labs(x = "Age in years", y="Probability of producing a narrative component")+
  theme(legend.position = c(0.9, 0.1), legend.background = element_rect(colour = "black", size = 0.5, linetype = "solid"))

```


```{r}
ggsave("../visuals/fig3.png", width = 10, height = 6, scale = 1)
```

# 2. How does the development of bilinguals’ narrative skills in L1 compare to that in L2? 

```{r}
mdata2 <- mdata%>%
  filter(ling == "Bilingual")%>%
  group_by(child, age, mode, testlanguage, test_lang_status)%>%
  summarise(sum = sum(response),
            n = length(response))


bm2 <- brm(sum|trials(n) ~ age + test_lang_status + mode + (1|child) + (age | testlanguage), 
    data = mdata2,
    chains = 4, 
    cores = 4,
    iter = 4000,
    family = binomial(),
    threads = threading(8), #to speed things up, comment out if not on a cluster
    backend = "cmdstanr",
    control = list(adapt_delta = 0.9, max_treedepth = 20)
    )%>%add_criterion("loo") 


saveRDS(bm2, "../saves/bm2.rds")

bm2_1 <- brm(sum|trials(n) ~ age * test_lang_status + mode + (1|child) + (age | testlanguage), 
    data = mdata2,
    chains = 4, 
    cores = 4,
    iter = 4000,
    family = binomial(),
    threads = threading(8), #to speed things up, comment out if not on a cluster
    backend = "cmdstanr",
    control = list(adapt_delta = 0.9, max_treedepth = 20)
    )%>%add_criterion("loo") #to speed things up, comment out if not on a cluster)


saveRDS(bm2_1, "../saves/bm2_1.rds")
```

## Model comparison

```{r}
print(loo_compare(bm2, bm2_1)%>%as_tibble(rownames = "model"))
```
model elpd_diff  se_diff    elpd_loo   se_elpd_loo p_loo  se_p_loo looic
  <chr> <compar.l> <compar.l> <compar.l> <compar.l>  <comp> <compar> <com>
1 bm2_1  0.000000  0.000000   -6647.887  46.74597    753.0… 20.35827 1329…
2 bm2   -1.292518  3.300637   -6649.180  46.64971    754.5… 20.28111 1329…


## Fixed effects

```{r}
fixef(bm2)
```

                     Estimate  Est.Error         Q2.5      Q97.5
Intercept           0.1688634 0.08472883 -0.001854431  0.3355942
age                 0.5224211 0.09062137  0.341886900  0.6984061
test_lang_statusL2  0.1792070 0.06218985  0.058240500  0.3041419
modetelling        -0.4113573 0.02331139 -0.455899700 -0.3655539

## Visualization

```{r}
# ndata2 <- mdata2%>%
#   select(testlanguage,test_lang_status, age, mode)%>%
#   group_by(testlanguage,test_lang_status, mode)%>%
#   distinct(age)%>%
#   mutate(n = 1)
# 
# 
# post2 <- fitted(bm2, newdata = ndata2, re_formula =~(age | testlanguage))%>%
#   as_tibble() %>%
#   bind_cols(ndata2)%>%
#   mutate(age = age * sd(data$age) + mean(data$age))%>%
#   mutate(mode = str_to_title(mode))%>%
#   mutate(mode = factor(mode, levels = c("Telling","Retelling")))
# 
# 
# ndata2_2 <- mdata2%>%
#   ungroup()%>%
#   distinct(mode, test_lang_status)%>%
#   expand_grid(age = unique(mdata2$age))%>%
#   mutate(n = 1)
# 
# post2_2 <- fitted(bm2, newdata = ndata2_2, re_formula = NA)%>%
#   as_tibble() %>%
#   bind_cols(ndata2_2)%>%
#   mutate(age = age * sd(data$age) + mean(data$age))%>%
#   mutate(mode = str_to_title(mode))%>%
#   mutate(mode = factor(mode, levels = c("Telling","Retelling")))
# 
# 
# saveRDS(post2, "../saves/post2.rds")
# saveRDS(post2_2, "../saves/post2_2.rds")


post2 <- readRDS("../saves/post2.rds")
post2_2 <- readRDS("../saves/post2_2.rds")


ggplot()+
  geom_line(data = post2, aes(x = age/12, y = Estimate, ymin = Q2.5, ymax = Q97.5, col = test_lang_status, group = interaction(test_lang_status,testlanguage)), stat = "identity", alpha = .25)+
  geom_smooth(data = post2_2, aes(x = age/12, y = Estimate, ymin = Q2.5, ymax = Q97.5, fill =test_lang_status, col = test_lang_status), stat = "identity", alpha = .25)+
  theme_bw()+
  scale_color_manual(name = "Test language", values = c("#345c83", "#40531B"))+
  scale_fill_manual(name = "Test language", values = c("#345c83", "#40531B"))+
  ylim(0,1)+
  facet_grid(~mode)+
  labs(x = "Age", y="Probability of producing a narrative component")+
  theme(legend.position = c(0.85, 0.15), legend.background = element_rect(colour = "black", size = 0.5, linetype = "solid"))
```

```{r}
ggsave("../visuals/fig4.png", width = 10, height = 4, scale = 1)
```

# 3. What is the general pattern in children’s use of factual and inferred components across the investigated languages?

run script "run_bm3.R"

```{r}
mdata_item <- data%>%
  mutate(age = scale(age))%>%
  group_by(child, age, ling, mode,testlanguage, item)%>%
  summarise(sum = sum(response),
            n = length(response))
```


```{r}
bm3 <- readRDS("../saves/bm3.rds")
```

## Fixed effects

```{r}
fixef(bm3)
```

                      Estimate  Est.Error       Q2.5       Q97.5
Intercept            0.5736588 0.66094813 -0.8164857  1.87133975
age                  0.7904910 0.28360935  0.2838487  1.33117925
lingMonolingual     -0.1295025 0.10735855 -0.3360821  0.08074064
modetelling         -0.4118338 0.02692149 -0.4640448 -0.35936258
age:lingMonolingual -0.5162402 0.14297884 -0.7930640 -0.23489268


## Item effects

```{r}
ranef(bm3)$item
```


          Estimate Est.Error       Q2.5     Q97.5
attempt  0.8244836 0.6541805 -0.4765082 2.2126055
goal    -0.8948328 0.6547706 -2.1964730 0.4923131
ist     -0.5591364 0.6529128 -1.8568003 0.8227929
outcome  0.8138825 0.6530311 -0.4749704 2.2059632

## Visualization

### Probability of component by language for avergae age
```{r}
ndata3 <- mdata%>%
  distinct(item)%>%
  expand_grid(testlanguage = unique(mdata$testlanguage))%>%
  mutate(mode = "telling", 
         age = 0, 
         ling = "Bilingual", 
         n = 1)


post3 <- fitted(bm3, newdata = ndata3, re_formula =~(age + item | testlanguage))%>%
  as_tibble() %>%
  bind_cols(ndata3)%>%
  mutate(testlanguage = recode(testlanguage,
                 "per" = "Persian",
                     "hrv" = "Croatian",
                     "rus" = "Russian",
                     "est" = "Estonian",
                     "fin" = "Finnish",
                     "swe" = "Swedish",
                     "geo" = "Georgian",
                     "tur" = "Turkish",
                     "aka" = "Akan",
                     "urd" = "Urdu",
                     "hin" = "Hindi",
                     "kan" = "Kannada",
                     "gon" = "Gondi",
                     "hlb" = "Halbi",
                     "gle" = "Irish",
                     "heb" = "Hebrew",
                     "ita" = "Italian",
                     "uzb" = "Uzbek",
                     "bi" = "Bislama",
                     "nld" = "Dutch",
                     "tgl" = "Tagalog",
                     "pol" = "Polish",
                     "eng" = "English",
                     "slk" = "Slovak",
                     "sgh" = "Shughni",
                     "lug" = "Luganda",
                     "sah" = "Sakha", 
                     "cor" = "Coriglianese",
                     "dan" = "Danish",
                     "bis" = "Bislama",
                     "iii" = "Sichuan Yi",
                     "cat" = "Catalan", 
                     "ita" = "Italian",
                     "ger" = "German",
                     "kaz" = "Kazakh",
                     "uig" = "Uighur",
                     "lt" = "Lithuanian",
                     "tgk" = "Tajik",
                     "msa" = "Malay",
                     "esp" = "Spanish",
                     "kor" = "Korean",
                     "cg" = "Cypriot Greek",
                     "yue" = "Yue Chinese",
                     "ell" = "Modern Greek",
                     "ber" = "Berber languages",
                     "ary" = "Moroccan Arabic",
                     "xho" = "Xhosa",
                     "afr" = "Afrikaans",
                     "cmn" = "Mandarin Chinese",
                     "ghc" = "Hiberno-Scottish Gaelic",
                     "ukr"  = "Ukrainian",
                      "lit" = "Lithuanian"
                           ))%>%
  filter(item != "setting")%>%
  mutate(item = recode(item,
                        "ist" = "Internal State", 
                       "goal" = "Goal",
                       "attempt" = "Attempt",
                       "outcome" = "Outcome"))%>%
  mutate(item = factor(item, levels = c("Internal State", "Goal", "Attempt", "Outcome")))


saveRDS(post3, "../saves/post3.rds")

post3 <- readRDS("../saves/post3.rds")

ggplot()+
  geom_pointrange(data = post3, aes(y = reorder(testlanguage, Estimate), x = Estimate, xmin = Q2.5, xmax = Q97.5, col = item), alpha = .75, position = position_dodge(width = 0.5))+
  theme_bw()+
  scale_color_viridis_d(name = "Component")+
  labs(y = "Language", x="Probability to produce the component")

```

```{r}
ggsave("../visuals/fig4.png", height = 8, width = 12, scale = 1)
```

### Probability of component by age across languages

```{r}
ndata3_1 <- mdata%>%
  distinct(item, age, ling, testlanguage)%>%
  mutate(mode = "telling", 
         n = 1)

post3_1 <- fitted(bm3, newdata = ndata3_1, re_formula =~(age + item | testlanguage), )%>%
  as_tibble() %>%
  bind_cols(ndata3_1)%>%
  mutate()%>%
  filter(item != "setting")%>%
  mutate(item = recode(item,
                        "ist" = "Internal State", 
                       "goal" = "Goal",
                       "attempt" = "Attempt",
                       "outcome" = "Outcome"))%>%
  mutate(item = factor(item, levels = c("Internal State", "Goal", "Attempt", "Outcome")))%>%
  mutate(age = age * sd(data$age) + mean(data$age))%>%
  mutate(mode = str_to_title(mode))

ndata3_2 <- mdata%>%
  distinct(item, age, ling)%>%
  mutate(mode = "telling", 
         n = 1)

post3_2 <- fitted(bm3, newdata = ndata3_2, re_formula = NA )%>%
  as_tibble() %>%
  bind_cols(ndata3_2)%>%
  mutate()%>%
  filter(item != "setting")%>%
  mutate(item = recode(item,
                        "ist" = "Internal State", 
                       "goal" = "Goal",
                       "attempt" = "Attempt",
                       "outcome" = "Outcome"))%>%
  mutate(item = factor(item, levels = c("Internal State", "Goal", "Attempt", "Outcome")))%>%
  mutate(age = age * sd(data$age) + mean(data$age))%>%
  mutate(mode = str_to_title(mode))


saveRDS(post3_2, "../saves/post3_2.rds")
saveRDS(post3_1, "../saves/post3_1.rds")

post3_1 <- readRDS("../saves/post3_1.rds")
post3_2 <- readRDS("../saves/post3_2.rds")

ggplot()+
   geom_line(data = post3_1, aes(x = age/12, y = Estimate, ymin = Q2.5, ymax = Q97.5, col = item, group = interaction(item,testlanguage)), stat = "identity", alpha = .15)+
  geom_smooth(data = post3_2, aes(x = age/12, y = Estimate, ymin = Q2.5, ymax = Q97.5, fill =item, col = item), stat = "identity", alpha = .25)+
  theme_bw()+
  scale_color_viridis_d(name = "Component")+
  scale_fill_viridis_d(name = "Component")+
  facet_grid(~ling)+
  labs(y = "Probability to produce the component", x="Age")+
  theme(legend.position = c(0.9, 0.2))
```

```{r}
ggsave("../visuals/fig5.png", width = 10, height = 5, scale = 1)
```

# 4. How do complete episodes evolve across languages and in monolingual vs bilingual children?

```{r}
# read in data coded for complete narratives (attempt, goal or internal state, and outcome)
nar_data1 <- read_csv("../data/nar_data1.csv")
```

```{r}
# mnar_data1 <- nar_data1%>%
#   mutate(age = scale(age))%>%
#   group_by(child, age, mode, testlanguage, ling)%>%
#   summarise(sum = sum(complete),
#             n = length(complete))
# 
# bm4 <- brm(sum|trials(n) ~ age + ling + mode + (1|child) + (age | testlanguage), 
#     data = mnar_data1,
#     chains = 4, 
#     cores = 4,
#     iter = 4000,
#     family = binomial(),
#     threads = threading(8), #to speed things up, comment out if not on a cluster
#     backend = "cmdstanr",
#     control = list(adapt_delta = 0.95, max_treedepth = 20)
#     )%>%add_criterion("loo")  #to speed things up, comment out if not on a cluster)
# 
# 
# saveRDS(bm4, "../saves/bm4.rds")
# 
# bm4_1 <- brm(sum|trials(n) ~ age * ling + mode + (1|child) + (age | testlanguage), 
#     data = mnar_data1,
#     chains = 4, 
#     cores = 4,
#     iter = 4000,
#     family = binomial(),
#     threads = threading(8), #to speed things up, comment out if not on a cluster
#     backend = "cmdstanr",
#     control = list(adapt_delta = 0.95, max_treedepth = 20)
#     )%>%add_criterion("loo") 
# 
# saveRDS(bm4_1, "../saves/bm4_1.rds")

```

## Model comparison

```{r}
#print(loo_compare(bm4, bm4_1))
```

      elpd_diff se_diff
bm4_1  0.0       0.0   
bm4   -2.6       2.8 

## Fixed effects

### Model wo interaction

```{r}
fixef(bm4)
```

                   Estimate  Est.Error       Q2.5      Q97.5
Intercept       -0.38729022 0.13845736 -0.6656797 -0.1214909
age              0.86070754 0.13449387  0.6122557  1.1350065
lingMonolingual -0.01216158 0.07802259 -0.1638015  0.1423151
modetelling     -0.70627605 0.04329577 -0.7927617 -0.6223730

### Interaction model 

```{r}
fixef(bm4_1)
```

                      Estimate  Est.Error       Q2.5       Q97.5
Intercept           -0.3435852 0.13694491 -0.6191936 -0.08773419
age                  0.9620635 0.15446200  0.6755273  1.28608100
lingMonolingual     -0.1224120 0.08678811 -0.2979445  0.04560172
modetelling         -0.7121734 0.04281527 -0.7960030 -0.62755342
age:lingMonolingual -0.3171017 0.12125556 -0.5532681 -0.07933007

### Visualization

```{r}
# ndata4 <- mnar_data1%>%
#   ungroup()%>%
#   distinct(testlanguage)%>%
#   expand_grid(ling = unique(mnar_data1$ling))%>%
#   mutate(mode = "telling", 
#          age = 0, 
#          n = 1)
# 
# 
# post4 <- fitted(bm4_1, newdata = ndata4, re_formula =~(1 | testlanguage))%>%
#   as_tibble() %>%
#   bind_cols(ndata4)%>%
#   mutate(testlanguage = recode(testlanguage,
#                  "per" = "Persian",
#                      "hrv" = "Croatian",
#                      "rus" = "Russian",
#                      "est" = "Estonian",
#                      "fin" = "Finnish",
#                      "swe" = "Swedish",
#                      "geo" = "Georgian",
#                      "tur" = "Turkish",
#                      "aka" = "Akan",
#                      "urd" = "Urdu",
#                      "hin" = "Hindi",
#                      "kan" = "Kannada",
#                      "gon" = "Gondi",
#                      "hlb" = "Halbi",
#                      "gle" = "Irish",
#                      "heb" = "Hebrew",
#                      "ita" = "Italian",
#                      "uzb" = "Uzbek",
#                      "bi" = "Bislama",
#                      "nld" = "Dutch",
#                      "tgl" = "Tagalog",
#                      "pol" = "Polish",
#                      "eng" = "English",
#                      "slk" = "Slovak",
#                      "sgh" = "Shughni",
#                      "lug" = "Luganda",
#                      "sah" = "Sakha", 
#                      "cor" = "Coriglianese",
#                      "dan" = "Danish",
#                      "bis" = "Bislama",
#                      "iii" = "Sichuan Yi",
#                      "cat" = "Catalan", 
#                      "ita" = "Italian",
#                      "ger" = "German",
#                      "kaz" = "Kazakh",
#                      "uig" = "Uighur",
#                      "lt" = "Lithuanian",
#                      "tgk" = "Tajik",
#                      "msa" = "Malay",
#                      "esp" = "Spanish",
#                      "kor" = "Korean",
#                      "cg" = "Cypriot Greek",
#                      "yue" = "Yue Chinese",
#                      "ell" = "Modern Greek",
#                      "ber" = "Berber languages",
#                      "ary" = "Moroccan Arabic",
#                      "xho" = "Xhosa",
#                      "afr" = "Afrikaans",
#                      "cmn" = "Mandarin Chinese",
#                      "ghc" = "Hiberno-Scottish Gaelic",
#                      "ukr"  = "Ukrainian"
#                            ))


saveRDS(post4, "../saves/post4.rds")

post4 <- readRDS("../saves/post4.rds")


ggplot()+
  geom_pointrange(data = post4, aes(y = testlanguage, x = Estimate, xmin = Q2.5, xmax = Q97.5, col = ling), alpha = .75, position = position_dodge(width = 0.5))+
  theme_bw()+
  scale_color_ptol(name = "Language status")+
  labs(y = "Language", x="Probability of producing complete narrative")+
  theme(legend.position = c(0.85, 0.5), legend.background = element_rect(colour = "black", size = 0.5, linetype = "solid"))
```

```{r}
ggsave("../visuals/fig7.png", height = 10, width = 8, scale = 1)
```

# 4.1. Complete narratives in bilinguals

```{r}
nar_data1_bi <- nar_data1%>%
  filter(ling == "Bilingual")%>%
  mutate(age = scale(age))%>%
  group_by(child, age, mode, testlanguage, test_lang_status)%>%
  summarise(sum = sum(complete),
            n = length(complete))

# bm4_bi <- brm(sum|trials(n) ~ age + test_lang_status + mode + (1|child) + (age | testlanguage), 
#     data = nar_data1_bi,
#     chains = 4, 
#     cores = 4,
#     iter = 4000,
#     family = binomial(),
#     threads = threading(8), #to speed things up, comment out if not on a cluster
#     backend = "cmdstanr",
#     control = list(adapt_delta = 0.95, max_treedepth = 20)
#     )%>%add_criterion("loo") 
# 
# saveRDS(bm4_bi, "../saves/bm4_bi.rds")

```

## Fixed effects

```{r}
fixef(bm4_bi)
```

                     Estimate  Est.Error       Q2.5       Q97.5
Intercept          -0.3285612 0.13023396 -0.5910429 -0.07849624
age                 0.8695058 0.14909338  0.5773645  1.17417125
test_lang_statusL2 -0.1023116 0.12358917 -0.3463284  0.14515982
modetelling        -0.5412475 0.05702476 -0.6532675 -0.43036353

# 4.2. Complete narratives ignoring internal states (narrative = goal + attempt + outcome)

```{r}
# read in data coded for complete narratives (attempt, goal, and outcome)
nar_data2 <- read_csv("../data/nar_data2.csv")
```

```{r}
mnar_data2 <- nar_data2%>%
  mutate(age = scale(age))%>%
  group_by(child, age, mode, testlanguage, ling)%>%
  summarise(sum = sum(complete),
            n = length(complete))

# bm4_2 <- brm(sum|trials(n) ~ age + ling + mode + (1|child) + (age | testlanguage), 
#     data = mnar_data2,
#     chains = 4, 
#     cores = 4,
#     iter = 4000,
#     family = binomial(),
#     threads = threading(8), #to speed things up, comment out if not on a cluster
#     backend = "cmdstanr",
#     control = list(adapt_delta = 0.95, max_treedepth = 20)
#     )%>%add_criterion("loo")  #to speed things up, comment out if not on a cluster)
# 
# 
# saveRDS(bm4_2, "../saves/bm4_2.rds")
```

## Fixed effects

```{r}
fixef(bm4_2)
```

                   Estimate  Est.Error       Q2.5      Q97.5
Intercept       -1.42736854 0.15386009 -1.7506915 -1.1394465
age              0.72225965 0.15177434  0.4478233  1.0475090
lingMonolingual -0.06789734 0.09495004 -0.2536275  0.1184465
modetelling     -0.80409980 0.05226774 -0.9060773 -0.7017779

# 5. Narrative elemtens depending on story
```{r}
storydata <- data%>%
  filter(!is.na(story))%>%
  mutate(story = ifelse(
    story == "cat" | story == "dog", "cat/dog", "baby birds/goats"
  ))%>%
  mutate(age = scale(age))%>%
  group_by(child, testlanguage, age, story, item, mode, ling)%>%
  summarise(sum = sum(response),
            n = length(response))

storydata <- data%>%
  filter(!is.na(story))%>%
  mutate(story = ifelse(
    story == "cat" | story == "dog", "cat/dog", "baby birds/goats"
  ))%>%
  mutate(age = scale(age))%>%
  group_by(child, testlanguage, age, story, item, mode, ling)%>%
  summarise(sum = sum(response),
            n = length(response))
```
run script `run_bm5.R`

```{r}
bm5 <- readRDS("../saves/bm5.rds")
bm5_1 <- readRDS("../saves/bm5_1.rds")
```

```{r}
print(loo_compare(bm5, bm5_1))
```

## Fixed effects

```{r}
round(fixef(bm5),2)
```

                             Estimate Est.Error  Q2.5 Q97.5
Intercept                        1.84      0.14  1.57  2.13
age                              0.70      0.12  0.47  0.95
lingMonolingual                 -0.13      0.05 -0.23 -0.03
modetelling                     -0.90      0.03 -0.95 -0.85
itemgoal                        -1.68      0.15 -1.98 -1.38
itemist                         -1.73      0.10 -1.93 -1.52
itemoutcome                     -0.14      0.13 -0.38  0.12
storycatDdog                    -1.15      0.11 -1.37 -0.93
age:lingMonolingual             -0.36      0.07 -0.49 -0.23
age:itemgoal                    -0.25      0.06 -0.37 -0.14
age:itemist                      0.15      0.05  0.05  0.25
age:itemoutcome                  0.06      0.06 -0.06  0.17
itemgoal:storycatDdog            0.82      0.07  0.69  0.95
itemist:storycatDdog             0.99      0.06  0.88  1.10
itemoutcome:storycatDdog         1.27      0.07  1.13  1.40
age:storycatDdog                 0.12      0.06  0.01  0.23
age:itemgoal:storycatDdog        0.09      0.06 -0.03  0.21
age:itemist:storycatDdog        -0.13      0.05 -0.23 -0.02
age:itemoutcome:storycatDdog     0.00      0.06 -0.12  0.12

## Visualization

### Probability of component by item and story
```{r}
ndata5 <- storydata%>%
  select(ling,item, story, testlanguage, age)%>%
  group_by(ling,item, story, testlanguage)%>%
  distinct(age, .keep_all = T)%>%
  mutate(mode = "telling", 
         n = 1)

post5 <- fitted(bm5, newdata = ndata5, re_formula =~(age + item + story | testlanguage))%>%
  as_tibble() %>%
  bind_cols(ndata5)%>%
  mutate()%>%
  filter(item != "setting")%>%
  mutate(item = recode(item,
                        "ist" = "Internal State", 
                       "goal" = "Goal",
                       "attempt" = "Attempt",
                       "outcome" = "Outcome"))%>%
  mutate(item = factor(item, levels = c("Internal State", "Goal", "Attempt", "Outcome")))%>%
  mutate(age = age * sd(data$age) + mean(data$age))%>%
  mutate(mode = str_to_title(mode))


ndata5_1 <- storydata%>%
  select(ling,item, story, age)%>%
  group_by(ling,item, story)%>%
  distinct(age, .keep_all = T)%>%
  mutate(mode = "telling", 
         n = 1)

post5_1 <- fitted(bm5, newdata = ndata5_1, re_formula = NA)%>%
  as_tibble() %>%
  bind_cols(ndata5_1)%>%
  mutate()%>%
  filter(item != "setting")%>%
  mutate(item = recode(item,
                        "ist" = "Internal State", 
                       "goal" = "Goal",
                       "attempt" = "Attempt",
                       "outcome" = "Outcome"))%>%
  mutate(item = factor(item, levels = c("Internal State", "Goal", "Attempt", "Outcome")))%>%
  mutate(age = age * sd(data$age) + mean(data$age))%>%
  mutate(mode = str_to_title(mode))

saveRDS(post5, "../saves/post5.rds")
saveRDS(post5_1, "../saves/post5_1.rds")

post5 <- readRDS("../saves/post5.rds")
post5_1 <- readRDS("../saves/post5_1.rds")

ggplot()+
  geom_line(data = post5, aes(x = age/12, y = Estimate, ymin = Q2.5, ymax = Q97.5, col = story, group = interaction(story,testlanguage)), stat = "identity", alpha = .25)+
  geom_smooth(data = post5_1, aes(x = age/12, y = Estimate, ymin = Q2.5, ymax = Q97.5, fill =story, col = story), stat = "identity", alpha = .25)+
  theme_bw()+
  scale_color_colorblind(name = "Story")+
  scale_fill_colorblind(name = "Story")+
  ylim(0,1)+
  facet_grid(ling~item)+
  labs(x = "Age", y="Probability of producing a narrative component")+
  theme(legend.position = c(0.1, 0.4), legend.background = element_rect(colour = "black", size = 0.5, linetype = "solid"))

```

```{r}
ggsave("../visuals/fig8.png", height = 6, width = 10, scale = 1)
```
